---
title: "midterm_key_oct_8"
author: "Rob Wells"
date: "2025-10-15"
output: html_document
name: Bella DeCrescenzo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam - Fall 2025

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

# Load tidyverse
```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade.

## Task 1: Read the data**
```{r}
inspections <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


## Task 2: Calculate the mean and standard deviation maximum and minimum of compliance scores. Put the results in a new dataframe called "compliance_stats." 
```{r}
compliance_stats <- inspections|>
    summarize(mean_score = mean(compliance_score),
          median_score = median(compliance_score),
          max_score = max (compliance_score),
          min_score = min (compliance_score),
          sd_score = sd(compliance_score))

```

## Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a headline. Include a headline and label the x and y axes.
```{r}
inspections |> 
  ggplot() + 
  geom_histogram(aes(x = compliance_score), binwidth = 1) +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", size = 1) +
  labs(title = "Distribution of Restaurants' Compliance Scores in Montgomery County", x = "Compliance Score", y = "Number of restaurants")



```

## Task 4: Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? 

The histogram and standard deviation tell us that the compliance scores are skewed left, which means that the median is greater than the mean. Therefore, 68% of the data falls between about 91 and 100 for their compliance scores. Therefore, the story is that the majority of restaurants in Montgomery County are highly compliant as the histogram is skewed left. 


## Part 2: High School Athletics (30 points)

The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and gender. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv`

# Task 5: Load the high school participation data
```{r}
participation <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```


#Task 6: Add two columns called total and girls_pct.  
Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. 
```{r}
participation_rates <- participation |>
  mutate(total = boys + girls) |>
  mutate(girls_pct = (girls/total) * 100)
```

#Task 7: Calculate correlation between boys and girls participation
```{r}
participation |> summarize(correlation = cor(boys, girls, method="pearson"))
```

#Task 8: Create a scatterplot showing this relationship, adding a line of best fit. Include a headline and label the x and y axes.
```{r}
participation |> 
  ggplot() +
  geom_point(aes(x=girls, y=boys)) +
  geom_smooth(aes(x=girls, y=boys), method="lm")+
  labs(title = "Relationship between gender and MD highschool sports participation", x = "Girls' participation in sports", y = "Boys' Participation in sports")



```


# Task 9: In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about gender equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

The correlation coefficient of about .98 and scatterplot reveals that there is almost a very strong positive correlation between girls' participation in high school sports in Maryland and boys' participation in high school sports in Maryland. This shows that gender equity in Maryland high school sports is very good and pretty equal, while some school districts that are above the line have a larger proportion of boys than girls participating and sports and school districts below the line have a larger proportion of girls participating in high school sports than boys. I would want to look into Baltimore County Public Schools, because their data is most disproportionate with less boys participating in sports in proportion to girls, which goes against societal norms and I would want to examine further.

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

#Task 10: Load the car thefts data
```{r}
car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

#Task 11: Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate 
```{r}
car_thefts_rates <- car_thefts |>
  mutate(car_thefts_per_county_2023 = (2023/population) * 10000)
```

#Task 12: Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median? 
```{r}
car_thefts_rates |>
  summarize(median_car_theft_rate = median(car_thefts_per_county_2023))
#Counties such as Dorchester County and Kent County have rates way above the median.
car_thefts |>
  summarize(total_car_thefts = sum(2023))
```


#Task 13: Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? 

These calculations, including the median car theft rate being 185, reveal that the distribution of car thefts in Maryland is skewed left as over half of the counties' car theft rates are above this median car theft rate. Therefore, the lede of a story could be: In Kent County, car theft rates are over 5 times larger than the median car theft rate throughout Maryland counties. 

## Part 4: Data Analysis Scenario (20 points)

#Task 14: Essay Question.

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
2. What visualizations would help readers understand the trends? (5 points)
3. What additional context or data would you need to make this a complete story? (5 points)

To verify this claim, I would calculate the mean response times per month, the median response time, and the standard deviation. A visualization such as a scatterplot would be most ideal in helping readers understand this trend. I would put response time on the x axis and the months on the y axis to show how these response times have actually changed over the past year. To complete this story, I would need to provide information about changes among the police and fire departments that might have contributed to this change.

# Excellent Response


### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Submit both your code and written responses
- Load this test to GitHub and submit the link in Elms

Good luck!
