---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

**A:**
-Our group's mean for colored M&M's was very far from the true proportion of .5 as we got 92.45%.
-Different groups get different results because we all have different sample sizes with randomly selected M&Ms.
-When we combine all groups the results will be more normal and closer to the true proportion of .5.
-This relates to polls adn surveys in realy life because we can see how differing sample sizes and sample methods impact poll and survey findings.

------------------------------------------------------------------------------

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**
The sample size was only about 12,000, and the response rate was only 23%. This means that the population size is way larger than the sample which causes me to raise questions of how accurate these findings are. I am also raising questions about how they included the 14 open response questions into the data. I don't think that having open response questions is a good way to get accurate, consistent results. They sent this survey out to the whole community and only got data from people that completed the survey.
------------------------------------------------------------------------------

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}
library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
younguns <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

head(younguns)

```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_100 <-younguns |> 
  sample_n(100)


```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_100 |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}
younguns |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics
**A**
```{r}
The sample and population statistics are almost exactly the same with the sample mean being 48.75, the population mean being 49.873, the sample sd being 18.2 and the population sd being 17.5.

```

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

**A**:
The means and standard deviations between the sample and population are very similar. When I made the sample size 20, the mean was 45.05 and the sd was 20.587, so I can see how the mean and sd vary more from the population mean and sd when a smaller sample size is taken. This shows how the smaller the sample size, the less accurate the statistics are from the sample data.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
