---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

From a reporting standpoint, I think that categories such as assault shootings, theft, and robbery are going to be the categories that citizens care about the most. The way that many of the categories are broken up definitely make the job of summarizing the data more difficult. For example, assaults are split up into three different categories, assault, assault shooting, and assault weapon. Therefore, while assaults decreased over the year, assault shootings increased. This makes reporting on assaults complicated because they are split up. It is also unclear if the assault category includes assault shootings and assault weapons.

### Task 4: Calculating Averages

Calculate the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime |>
  mutate(total_incidents = `Jan`+`Feb`+`Mar`+`Apr`+`May`+`Jun`+`Jul`+`Aug`+`Sep`+`Oct`+`Nov`+`Dec`) |>
  mutate(average_incidents = total_incidents/12)
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

Accidents have the highest average number of incidents per month and the breaking and entering other category has the lowest average number of incidents per month. Within Prince George's County, crime seems to be centered around automobiles as the categories with the highest average number of incidents per month are auto stolen, accident, accident with impound, and theft from auto. I wonder what breaking and entering incidents are in the other category.

### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |>
  mutate(june_change = (average_incidents-`Jun`)/`Jun`*100) |>
mutate(july_change = (average_incidents-`Jul`)/`Jul`*100) |>
mutate(august_change = (average_incidents-`Aug`)/`Aug`*100)

```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

There are eleven categories that saw an increase from the average number of incidents in June. Then in July, there were eight categories that experienced an increase, and in August there were 7. Therefore, I saw an overall larger rise in crime in June than in July and August. While it can be seen that there was an overall increase in crime from the overall averages in the summer months, I would need to compare multiple years of data to establish this trend, as this sample size is too small to fully legitimize the theory.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Incidents of commercial breaking and entering in Prince George's County rose drastically in 2024. 

The number of commercial breaking and enterings per month in 2024 increased throughout the year, with an average of 37 commercial breaking and enterings for the first six months of the year and an average of 52 breaking and enterings for the last six months. There was even a 108% increase between the number of commercial breaking and enterings between data from January and December of 2024. 

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
