---
title: "HW3 WH Salaries"
name: Bella DeCrescenzo
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

What stands out to me is that there are several employees that say they make $0. With these people being directors and advisors, I don't see why they would be making $0 so this is definitely something I want to look into. This will also cause problems for analysis as it skews the data and messes with the mean/median.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
  wh_salaries |>
summarize(mean_salary= mean(salary),
    median_salary = median(salary))
```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |>
  summarize(mean_salary = mean(salary),
          median_salary = median(salary),
          max_salary = max (salary),
          min_salary = min (salary)
)
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

When I try to run the above code, I just get two warnings and it will not spit out the data. I believe my code is accurate so I'm not sure why I'm getting warnings and can't see the results of my analysis. However, with the minimum salary being zero which serves as an outlier, it is obvious that the data is going to be skewed and the mean/median are going to be affected. 

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

My story idea is looking into the employees who are said to make a salary of $0. This cannot be completely accurate and I want to get to the bottom of why it is included in this data as $0 when it says that they are employees. To develop this story idea, I would start looking into the figures that are said to make $0 and even contacting them to set up an interview to ask some questions.
