---
title: "final_project"
output: html_document
date: "2025-11-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
tuition <- read_csv("https://raw.githubusercontent.com/wellsdata/Jour405_student_repo_fall_2025/refs/heads/main/project/college_scorecard_2021_2024.csv")
```
***This dataset consists of the in-state and out-of-state tuitions in Maryland and Virginia colleges and universities from 2021 to 2023. The data includes the name of the institution, the tuition in-state and out-of-state tuitions from 2021-2023, and then the insititution's enrollment count per year. The data came from the U.S. Department of Education's College Scorecard. (https://collegescorecard.ed.gov/)

The most important columns for my analysis will be the in-state and out-of-state tuition columns for each year, as I will probably be looking at how these figures have changed over time. 

Something that I might want that is not included in the data is how inflation in the United States has changed throughout the years included in the dataset. Understanding how inflation has changed from 2021 to 2023 will allow me to more accurately understand tuition changes, especially increases, throughout the years.

```{r}
tuitions2021_analysis_instate <- tuition |>
  summarize(
    mean_2021_intuition = mean(`2021_22_tuitionfee_in`, na.rm=TRUE), 
    median_2021_intuition = median(`2021_22_tuitionfee_in`, na.rm=TRUE),
    mean_2021_intuition = mean(`2021_22_tuitionfee_in`, na.rm=TRUE), 
    max_2021_intuition = max(`2021_22_tuitionfee_in`, na.rm=TRUE),
    mean_2021_intuition = mean(`2021_22_tuitionfee_in`, na.rm=TRUE), 
    min_2021_intuition = min(`2021_22_tuitionfee_in`, na.rm=TRUE))
```

```{r}
tuitions2021_analysis_outofstate <- tuition |>
  summarize(
    mean_2021_outtuition = mean(`2021_22_tuitionfee_out`, na.rm=TRUE), 
    median_2021_outtuition = median(`2021_22_tuitionfee_out`, na.rm=TRUE),
    mean_2021_outtuition = mean(`2021_22_tuitionfee_out`, na.rm=TRUE), 
    max_2021_outtuition = max(`2021_22_tuitionfee_out`, na.rm=TRUE),
    mean_2021_outtuition = mean(`2021_22_tuitionfee_out`, na.rm=TRUE), 
    min_2021_outtuition = min(`2021_22_tuitionfee_out`, na.rm=TRUE))
```

Possible hypothesis:
Null: Tuition for Maryland and Virginia colleges and universities have not changed significantly from 2021-2023 for in-state and out-of-state students.
Alternative: Tuition for Maryland and Virginia colleges and universities have changed significantly from 2021-2023 for in-state and out-of-state students.

Null: Out-of-state students throughout Maryland and Virginia colleges and universities have not experienced a more significant change in tuition than in-state students from 2021-2023.
Alternative: Out-of-state students throughout Maryland and Virginia colleges and universities have experienced a more significant change in tuition than in-state students from 2021-2023.